---
title: "Lab 7: Birth Ratios"
author: "Nicole Amestegui"
date: "`r Sys.Date()`"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    highlight: tango
    toc: false
    df_print: kable
    fig_caption: no
    number_sections: no
    dev: pdf
  html_document:
    highlight: tango
    self_contained: true
    theme: paper
    toc: no
    df_print: kable
    fig_caption: no
    number_sections: no
    smart: yes
    dev: svg
  github_document:
    html_preview: false
    toc: true
    toc_depth: 4
    dev: svg
    df_print: kable
---

```{r setup, include = FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
# Set knitr options
knitr::opts_chunk$set(
  echo = TRUE, eval = TRUE, fig.width = 5,
  fig.asp = 0.618, out.width = "80%", dpi = 120,
  fig.align = "center", cache = FALSE
)
# Load packages
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(infer))
suppressPackageStartupMessages(library(HistData))
set.seed(42)
```

---

## Visualizing and quantifying the distribution

### Exercise 1

```{r}
Arbuthnot_dedup <- Arbuthnot %>%
  filter(Year != 1704)

dim(Arbuthnot_dedup)
```


### Exercise 2

```{r}
ggplot(data = Arbuthnot_dedup) +
  geom_boxplot(aes(x = "", y = Ratio)) +
  coord_flip()
```
```{r}
ggplot(data = Arbuthnot_dedup,
       aes(x = Ratio, y = after_stat(density))) +
  geom_histogram(binwidth = 0.01)
```
center- the distribution is centered around 1.07
shape - the distibution is slightly right skewed
spread- most ratios fall between abt 1.01 and 1.15
### Exercise 3

```{r}
ggplot(data = Arbuthnot_dedup, aes(x = Ratio)) +
  geom_histogram(aes(y = after_stat(density)),
                 binwidth = 0.01,
                 alpha = 0.5) +
  geom_density()
```
```{r}
Arbuthnot <- Arbuthnot %>%
  mutate(Total = Males + Females)
```



### Exercise 4
```{r}
Arbuthnot_dedup %>%
  summarize(
    mean = mean(Ratio),
    median = median(Ratio),
    sd = sd(Ratio),
    iqr = IQR(Ratio),
    min = min(Ratio),
    max = max(Ratio)
  )
```

the mean, standard deviation, minimum, and maximum change more with outliers.
## `infer`ing a trend

### Exercise 5

null hypothes - the mean male to female birth ratio is 1.
alternative hypothesis - the mean male to female birth ratio is not 1. 


### Exercise 6
```{r}
Arbuthnot_dedup_null_105 <- Arbuthnot_dedup %>%
  specify(formula = Ratio ~ NULL) %>%
  hypothesize(null = "point", mu = 1.05) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "mean")
```

```{r}
Arbuthnot_dedup_obs_stat <- Arbuthnot_dedup %>%
  specify(Ratio ~ NULL) %>%
  calculate(stat = "mean")
```

```{r}
Arbuthnot_dedup_null_105 %>%
  get_p_value(
    obs_stat = Arbuthnot_dedup_obs_stat,
    direction = "two_sided"
  )
```
```{r}
visualize(Arbuthnot_dedup_null_105) +
  shade_p_value(
    obs_stat = Arbuthnot_dedup_obs_stat,
    direction = "two_sided"
  )
```
Yes, the result is different because the null hypothesis changed from 1 to 1.05. However, the observed mean is still much larger than 1.05, so the null hypothesis is rejected

